Turbulence and Particle Acceleration in Giant Radio Haloes: the Origin of Seed Electrons

### 2016-08-31 paper.tex
unified comments to \C{}

\usepackage[normalem]{ulem}
\usepackage[usenames]{xcolor}
\newcommand\cp[1]{{\color{red} #1}}
\newcommand\po[1]{{\color{blue} #1}}
\newcommand\AP[1]{{\color{green} #1}}

@@

\begin{abstract}
  About one third of X-ray-luminous clusters show smooth, unpolarised radio
  emission on $\sim$Mpc scales, known as giant radio haloes. One promising model
  for radio haloes is Fermi II acceleration of seed relativistic electrons by
  compressible turbulence in the intracluster medium (ICM). The origin of these
  seed electrons has never been fully explored. Here, we integrate the
  Fokker-Planck equation of the cosmic ray (CR) electron and proton
  distributions in \AP{the post-processing of} cosmological simulations of
  cluster formation, and confront them with radio surface brightness and
  spectral data of Coma. For standard assumptions, structure formation
  shocks lead to a seed electron population which produces too centrally
  concentrated radio emission. By modifying properties of the CR population
  (rapid streaming; enhanced CR electron acceleration at shocks) or turbulence
  (increasing turbulent-to-thermal energy density with radius), we can match
  observations, but only at the expense of fine-tuning. We then perform a
  parameter study. Under naive assumptions, radio properties are exponentially
  sensitive to the amplitude of turbulence, which is inconsistent with small
  scatter in scaling relations. This sensitivity can be removed if we relate the
  acceleration time to the turbulent dissipation time. Their ratio depends only
  on plasma parameters. In this case, turbulence above a threshold value
  required to overcome cooling provides a fixed amount of amplification;
  observations could then potentially constrain the unknown CR seed
  population. \po{Given that MHD turbulence only begins at the Alfv{\'e}n scale, in
    order to have sufficient acceleration, the turbulent cascade has to
    terminate by transit time damping on CRs, i.e., thermal particles must be
    scattered by plasma instabilities.} Understanding the small scatter in radio
  halo scaling relations may prove a rich source of insight on plasma processes
  in the ICM.
\end{abstract}

@@

Given a seed population of CRs, we adopt essentially the same set of
plasma physics assumptions as the reacceleration model for RHs
\citep{brunetti07,brunetti11}. We solve the isotropic, gyro-phase
averaged Fokker-Planck equation (via a Crank-Nicholson scheme) for the
time evolution of the CRe distribution in the Lagrangian frame
\citep{brunetti07,brunetti11}:
\begin{eqnarray}
{{d f_{\rmn{e}}(p,t)}\over{d t}} &\!=&
\frac{\partial}{\partial p}
\left\{
f_{\rmn{e}}(p,t)\left[
\left|\frac{d p}{d t}\right|_{\rm Coul} 
+ \frac{p}{3}\left(\bnabla\bcdot \bvel\right)
+ \left|\frac{d p}{d t}\right|_{\rm rad}\right.\right.
\nonumber\\
&-& \left.\left.{1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right) 
\right]\right\} - \left(\bnabla\bcdot \bvel\right) f_{\rmn{e}}(p,t)
\nonumber\\
&+& {{\partial^2 }\over{\partial p^2}}
\left[
D_{pp} f_{\rmn{e}}(p,t) \right]+ Q_{\rmn{e}}\left[p,t;f_{\rmn{p}}(p,t)\right]   \, .
\label{elettroni}
\end{eqnarray}
Here $f_{\rmn{e}}$ is the one-dimensional distribution in position $x$
(suppressed for clarity), momentum \cp{$p=P/(m_{\rmn{e}} c)$} and time $t$
(which is normalised such that the number density is given by
$n_{\rmn{CRe}}(t)=\int d p f_{\rmn{e}}(p,t)$), $d/d t=\partial/\partial
t+\bvel\bcdot\bnabla$ is the Lagrangian derivative, $\bvel$ is the gas velocity,
$|d p/d t|$ represents Coulomb \citep[Coul,][]{1972Phy....60..145G} and
radiative \citep[rad,][]{1979rpa..book.....R} losses, respectively,
\begin{eqnarray}
   \left|\frac{d p}{d t}\right|_{\rm Coul} &=& \frac{3\,\sigma_\rmn{T}\,n_{\rmn{e}}\,c}{2\, \beta_{\rmn{e}}^2}
  \left[\ln\left(\frac{m_{\rmn{e}} c^2 \beta_{\rmn{e}} \sqrt{\gamma-1}}{\hbar\,\omega_\rmn{plasma}}\right)\right.\nonumber \\
    &-&\ln(2)\left(\frac{\beta_{\rmn{e}}^2}{2}+\frac{1}{\gamma}\right)+\frac{1}{2}+
    \left.\left(\frac{\gamma-1}{4\gamma}\right)^2\right]\,, \\
  \left|\frac{d p}{d t}\right|_{\rm rad~} &=& \frac{4}{3}\frac{\sigma_\rmn{T}}{m_{\rmn{e}}c}\,
  \frac{p^2}{\beta_{\rmn{e}}}\,\left[1+\left(\frac{B}{B_\rmn{CMB}}\right)^2\right]\,\cp{\eps_{\rmn{CMB}}}\,.
\end{eqnarray}
Here $\beta_{\rmn{e}} = p/\sqrt{1+p^2}$ is the dimensionless velocity of CRs,
$\gamma=\sqrt{1+p^2}$ is the Lorentz factor of CRs,
$\omega_\rmn{plasma} = \sqrt{4\upi e^2 n_{\rmn{e}} / m_{\rmn{e}}}$ is the plasma
frequency, $n_e$ is the number density of free electrons, and
$\sigma_\rmn{T}= 8\upi e^4/3(m_{\rmn{e}} c^2)^2$ is the Thomson cross
section. The {\it rms} magnetic field strength is denoted by $B$ and
the equivalent field strength of the cosmic-microwave background is
given by $B_\rmn{CMB} = 3.24 (1 + z)^2\umu\rmn{G}$, where $z$ denotes
the redshift. In the peripheral cluster regions, where $B \ll
B_\rmn{CMB}$, the CRes loose virtually all their energy by means of
inverse Compton emission. $D_{pp}$ is the momentum space diffusion
coefficient (see Section~\ref{sec:reacc}), and
$Q_{\rmn{e}}$ denotes the injection rate of primary and secondary
electrons in the ICM (see section~\ref{sec:cosmo_sim}). The first term
containing the expression $\bnabla\bcdot \bvel$ represents Fermi-I
acceleration and the second term of this form describes adiabatic
gains and losses.

During post-processing of our Coma-like cluster simulation, we solve the
Fokker-Planck equation over a redshift interval from $z=5$ to 0. The simulated
cluster undergoes a major merger over the last 1-2~Gyrs that is thought to
inject large turbulent eddies. As is commonly assumed
\citep{brunetti07,brunetti11,2004ApJ...614..757Y,2013ApJ...771..131B} we assume
that about one Gyr after core passage the fields have decayed down to the
smallest scales $k_{\rm cut}$, and the radio halo turns on shortly after. We
choose this simulation snapshot to analyse. \AP{Note that in recent simulations by \citet{miniati15}, the turbulent reacceleration is strongest around core passage. However, we are not very sensitive to the adopted decay time, since the thermal and CR quantities are very similar a few 100 Myrs before and after $z = 0$, where we have chosen to evaluate the simulations.}
In all our calculations we assume that turbulent reacceleration efficiently
accelerates particles for $\tau_{\rm cl} \sim 650$ Myrs (which is roughly the
cascade time on which turbulence is damped) and that during this turbulent phase
CR streaming and spatial diffusion can be neglected. In
Section~\ref{sect:self-limiting}, we explore sensitivity to the last assumption.

@@

The time evolution of the spectral energy distribution of CRps,
$f_{\rmn{p}}$, is similarly given by:
\begin{eqnarray}
\lefteqn{
  {{d f_{\rmn{p}}(p,t)}\over{d t}} =
  {{\partial }\over{\partial p}}
  \left\{
  f_{\rmn{p}}(p,t)\left[ \left|{{d p}\over{d t}}\right|_{\rm Coul}
    + \frac{p}{3}\bnabla\bcdot \left(\bvel+\bvel_{\rmn{st}}\right)\right.\right.}
\nonumber\\
&&-\left.\left. {1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right)
\right]\right\} - f_{\rmn{p}}(p,t) \bnabla\bcdot \left(\bvel+\bvel_{\rmn{st}}\right) 
- \bvel_{\rmn{st}}\bcdot\bnabla f_{\rmn{p}}(p,t)
\nonumber\\
&&+\,\,\, {{\partial^2 }\over{\partial p^2}}
\left[ D_{pp} f_{\rmn{p}}(p,t) \right] - {{f_{\rmn{p}}(p,t)}\over{\tau_{\rm had}(p)}}
+ Q_{\rmn{p}}(p,t)\, ,
\label{eq:FP_p}
\end{eqnarray}
where $\bvel_{\rmn{st}}=-\varv_{\rmn{A}}\bnabla f_{\rmn{p}}/|\bnabla
f_{\rmn{p}}|$ is the streaming velocity in the isotropic transport
approximation, $\varv_{\rmn{A}}$ is the Alfv\'en speed,
and $Q_{\rmn{p}}(p,t)$ denotes the injection rate of shock accelerated
CRps as a function of momentum \cp{$p=P/(m_{\rmn{p}} c)$} and time $t$ (see
section~\ref{sec:cosmo_sim}). The timescale of
hadronic losses that produce pions via CRp collisions with thermal
protons of the ICM is given by 
\begin{eqnarray}
  \tau_{\rm had} = \left[c\,n_\rmn{th}\,\sigma^{+/-,0}(p)\right]^{-1}\,,
\end{eqnarray}
where we use the cross-section, $\sigma^{+/-,0}(p)$, given by the
fitting formula in \citet{1986ApJ...307...47D}.

@@

We adopt a Kraichnan spectrum ($s=3/2$) for the fast modes, as seen in simulations \citep{cho03}. This can also be written as: 
\begin{equation}
  \label{eq:Wk}
  \mathcal{W}(k) =
\sqrt{2/7\,I_L\,\rho\,\langle \Vph \rangle}\,k^{-3/2},
\end{equation}
where $I_L$ is the volumetric injection rate of turbulence at scale $L$ (which
is assumed to be constant), $\varv_{\rm ph}\approx c_{\rm s}$ is the phase speed
of waves. \cp{Since the Kraichnan spectrum is critical in what follows, it is
  worthwhile taking a moment to recount the origin of the spectral slope $s$ and
  the cascade rate. In the magnetically dominated regime, two counterpropagating
  wave packets of scale $l$ interact on a wave crossing time $\tau_{\rm
    ph}=l/\varv_{\rm ph}$, rather than the eddy turnover time $\tau_{\rm edd} =
  l/\varv_{\rm l}$.  Since $\tau_{\rm ph} \ll \tau_{\rm edd}$, each interaction
  results in a small velocity change $\delta \varv_{\rm l} \sim \varv_{\rm l}
  (\tau_{\rm ph}/\tau_{\rm edd})$. If these changes behave like a random walk,
  the cascade time $\tau_{\rm l}$ it takes for an eddy to become non-linear
  ($\Delta \varv \sim \varv_{\rm l})$ and cascades to smaller scales occurs with
  a characteristic velocity $\varv_{\rm l} \sim \delta \varv_{\rm l} (\tau_{\rm
    l}/\tau_{\rm ph})^{1/2} \sim \varv_{\rm l} (\tau_{\rm ph}/\tau_{\rm edd})\,
  (\tau_{\rm l}/\tau_{\rm ph})^{1/2}$, which we can solve to obtain the cascade
  time:
\begin{equation}
  \label{eqn:tau_l}
\tau_{\rm l} \sim \left(\frac{\tau_{\rm edd}}{\tau_{\rm ph}}\right)^{2} \tau_{\rm ph} \sim \frac{l \varv_{\rm ph}}{\varv_{\rm l}^{2}} 
\end{equation}
where we have implicitly assumed isotropy. This implies a dissipation rate $\epsilon \sim \varv_{\rm l}^{2}/\tau_{\rm l} \sim \varv_{\rm l}^{4}/(l \varv_{\rm ph})$, or $\varv_{\rm l} \sim (\epsilon l \varv_{\rm ph})^{1/4}$. Thus, since $k E(k)\sim \varv_{\rm l}^{2}$, this gives a kinetic energy spectrum: 
\begin{equation}
E(k) \sim (\epsilon \varv_{\rm ph})^{1/2} k^{-3/2}.
\end{equation}}

It is important to realise that this Kraichnan scaling only applies at small scales, where turbulence is magneto-hydrodynamical. At large scales, turbulent motions are subsonic (${\mathcal M}_{\rm s} \sim 0.2-0.5$) but super-Alfv{\'e}nic (${\mathcal M}_{\rm A} \sim 5$). Thus, except in the case where motions are transsonic and weak shocks become important, motions are fundamentally hydrodynamic and turbulence follows a Kolmogorov ($s=5/3$) spectrum. However, while the turbulent energy density $\eps_{\rm turb} \propto l^{2/3}$ decreases at small scales, the magnetic energy density \po{(which is dominated by the large scale mean field)} is scale-independent. Thus, at some scale $l_{\rm A} \sim l_L {\mathcal M}_{\rm A}^{-3}$, where $\eps_{\rm turb} \sim \eps_{B}$, the magnetic field becomes dynamically important, and turbulence transitions to the MHD regime. The cutoff scale in equation (\ref{eqn:Wk}) can be computed by setting the turbulent cascade rate for fast modes to the transit time damping rate on thermal electrons. This yields \citep{brunetti07,miniati15}:
\begin{equation}
k_{\rm cut} \approx A k_{\rmn{outer}} \frac{\langle (\delta \varv_{\rmn{outer}})^{2}\rangle^{2}}{c_{\rm s}^{4}}
\label{eqn:k_c1} 
\end{equation}
where $A \approx 11 000$. \cp{If there is an unimpeded cascade from the injection to the cutoff scale, $k_{\rmn{outer}}=k_L$ and $\varv_{\rmn{outer}}=\varv_{\rmn{c}}$. However, in our case the hydrodynamic (Kolmogorov) cascade transitions to the MHD cascade at the Alfv{\'e}n scale so that $k_{\rmn{outer}}=k_{\rmn{A}}$ and $\varv_{\rmn{outer}}=\varv_{\rmn{A}}$ and hence, we get\footnote{\cp{Note that this differs from previous work, which assumes Kraichnan turbulence from the injection scale onward and effectively adopts wave number and compressible velocity at the outer scale $L$ for this estimate of $k_{\rm cut}$.}}
\begin{equation}
k_{\rm cut} \approx A k_{\rmn{A}} \beta^{-2}\,.
\label{eqn:k_c} 
\end{equation}
}
This gives $2\upi/k_\rmn{cut}\sim 0.1-1$~kpc in the ICM. This constitutes an
effective mean free path for CRs, unless plasma instabilities can
mediate interactions between turbulence and particles on smaller
scales \citep{brunetti11}, a possibility we discuss in Section~\ref{sect:self-limiting}. Another possibility is that compressible modes dissipate in weak shocks, resulting in Burgers' turbulence $s=2$ \citep{kowal10,porter15,miniati15}. If Burgers turbulence dominates, then particle acceleration rates are too slow in the face of cooling processes to explain radio haloes \citep{miniati15}, and an alternative model for radio haloes is required.

@@

What is a typical energy density in compressive fast modes? The total turbulent
energy is typically $\sim 15-70$ per cent of the thermal energy in a cluster
\citep{vazza11}; it rises rapidly during a merger. The compressible component is
$\sim 20-40$ per cent of the total turbulent energy, and shows more rapid temporal
variations compared to the \po{incompressible} component
\citep{2013ApJ...771..131B,miniati15}. Note the compressible component in
cluster simulations is much larger than in stirring box simulations
with similar Mach numbers \citep{kowal10,lynn14}. This is likely due to the
compressive nature of turbulent driving (transsonic infall and merger), whereas
idealised simulations tend to use incompressible solenoidal driving and allow
compressive fluctuations to develop on their own. Overall, we adopt a
compressive energy density which is $X_{\rm tu} \sim 0.2$ of the thermal energy
as a baseline estimate.

>>>>

\usepackage[normalem]{ulem}
\usepackage[usenames]{xcolor}
\newcommand\cp[1]{{\color{red} #1}}
\newcommand\po[1]{{\color{blue} #1}}
\newcommand\AP[1]{{\color{green} #1}}
\newcommand\C[1]{{\bf #1}}

@@

\begin{abstract}
  About one third of X-ray-luminous clusters show smooth, unpolarised radio
  emission on $\sim$Mpc scales, known as giant radio haloes. One promising model
  for radio haloes is Fermi II acceleration of seed relativistic electrons by
  compressible turbulence in the intracluster medium (ICM). The origin of these
  seed electrons has never been fully explored. Here, we integrate the
  Fokker-Planck equation of the cosmic ray (CR) electron and proton
  distributions in \C{the post-processing of} cosmological simulations of
  cluster formation, and confront them with radio surface brightness and
  spectral data of Coma. For standard assumptions, structure formation
  shocks lead to a seed electron population which produces too centrally
  concentrated radio emission. By modifying properties of the CR population
  (rapid streaming; enhanced CR electron acceleration at shocks) or turbulence
  (increasing turbulent-to-thermal energy density with radius), we can match
  observations, but only at the expense of fine-tuning. We then perform a
  parameter study. Under naive assumptions, radio properties are exponentially
  sensitive to the amplitude of turbulence, which is inconsistent with small
  scatter in scaling relations. This sensitivity can be removed if we relate the
  acceleration time to the turbulent dissipation time. Their ratio depends only
  on plasma parameters. In this case, turbulence above a threshold value
  required to overcome cooling provides a fixed amount of amplification;
  observations could then potentially constrain the unknown CR seed
  population. \C{Given that MHD turbulence only begins at the Alfv{\'e}n scale, in
    order to have sufficient acceleration, the turbulent cascade has to
    terminate by transit time damping on CRs, i.e., thermal particles must be
    scattered by plasma instabilities.} Understanding the small scatter in radio
  halo scaling relations may prove a rich source of insight on plasma processes
  in the ICM.
\end{abstract}

@@

Given a seed population of CRs, we adopt essentially the same set of
plasma physics assumptions as the reacceleration model for RHs
\citep{brunetti07,brunetti11}. We solve the isotropic, gyro-phase
averaged Fokker-Planck equation (via a Crank-Nicholson scheme) for the
time evolution of the CRe distribution in the Lagrangian frame
\citep{brunetti07,brunetti11}:
\begin{eqnarray}
{{d f_{\rmn{e}}(p,t)}\over{d t}} &\!=&
\frac{\partial}{\partial p}
\left\{
f_{\rmn{e}}(p,t)\left[
\left|\frac{d p}{d t}\right|_{\rm Coul} 
+ \frac{p}{3}\left(\bnabla\bcdot \bvel\right)
+ \left|\frac{d p}{d t}\right|_{\rm rad}\right.\right.
\nonumber\\
&-& \left.\left.{1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right) 
\right]\right\} - \left(\bnabla\bcdot \bvel\right) f_{\rmn{e}}(p,t)
\nonumber\\
&+& {{\partial^2 }\over{\partial p^2}}
\left[
D_{pp} f_{\rmn{e}}(p,t) \right]+ Q_{\rmn{e}}\left[p,t;f_{\rmn{p}}(p,t)\right]   \, .
\label{elettroni}
\end{eqnarray}
Here $f_{\rmn{e}}$ is the one-dimensional distribution in position $x$
(suppressed for clarity), momentum $p=P/(m_{\rmn{e}} c)$ and time $t$
(which is normalised such that the number density is given by
$n_{\rmn{CRe}}(t)=\int d p f_{\rmn{e}}(p,t)$), $d/d t=\partial/\partial
t+\bvel\bcdot\bnabla$ is the Lagrangian derivative, $\bvel$ is the gas velocity,
$|d p/d t|$ represents Coulomb \citep[Coul,][]{1972Phy....60..145G} and
radiative \citep[rad,][]{1979rpa..book.....R} losses, respectively,
\begin{eqnarray}
   \left|\frac{d p}{d t}\right|_{\rm Coul} &=& \frac{3\,\sigma_\rmn{T}\,n_{\rmn{e}}\,c}{2\, \beta_{\rmn{e}}^2}
  \left[\ln\left(\frac{m_{\rmn{e}} c^2 \beta_{\rmn{e}} \sqrt{\gamma-1}}{\hbar\,\omega_\rmn{plasma}}\right)\right.\nonumber \\
    &-&\ln(2)\left(\frac{\beta_{\rmn{e}}^2}{2}+\frac{1}{\gamma}\right)+\frac{1}{2}+
    \left.\left(\frac{\gamma-1}{4\gamma}\right)^2\right]\,, \\
  \left|\frac{d p}{d t}\right|_{\rm rad~} &=& \frac{4}{3}\frac{\sigma_\rmn{T}}{m_{\rmn{e}}c}\,
  \frac{p^2}{\beta_{\rmn{e}}}\,\left[1+\left(\frac{B}{B_\rmn{CMB}}\right)^2\right]\,\eps_{\rmn{CMB}}\,.
\end{eqnarray}
Here $\beta_{\rmn{e}} = p/\sqrt{1+p^2}$ is the dimensionless velocity of CRs,
$\gamma=\sqrt{1+p^2}$ is the Lorentz factor of CRs,
$\omega_\rmn{plasma} = \sqrt{4\upi e^2 n_{\rmn{e}} / m_{\rmn{e}}}$ is the plasma
frequency, $n_e$ is the number density of free electrons, and
$\sigma_\rmn{T}= 8\upi e^4/3(m_{\rmn{e}} c^2)^2$ is the Thomson cross
section. The {\it rms} magnetic field strength is denoted by $B$ and
the equivalent field strength of the cosmic-microwave background is
given by $B_\rmn{CMB} = 3.24 (1 + z)^2\umu\rmn{G}$, where $z$ denotes
the redshift. In the peripheral cluster regions, where $B \ll
B_\rmn{CMB}$, the CRes loose virtually all their energy by means of
inverse Compton emission. $D_{pp}$ is the momentum space diffusion
coefficient (see Section~\ref{sec:reacc}), and
$Q_{\rmn{e}}$ denotes the injection rate of primary and secondary
electrons in the ICM (see section~\ref{sec:cosmo_sim}). The first term
containing the expression $\bnabla\bcdot \bvel$ represents Fermi-I
acceleration and the second term of this form describes adiabatic
gains and losses.

During post-processing of our Coma-like cluster simulation, we solve the
Fokker-Planck equation over a redshift interval from $z=5$ to 0. The simulated
cluster undergoes a major merger over the last 1-2~Gyrs that is thought to
inject large turbulent eddies. As is commonly assumed
\citep{brunetti07,brunetti11,2004ApJ...614..757Y,2013ApJ...771..131B} we assume
that about one Gyr after core passage the fields have decayed down to the
smallest scales $k_{\rm cut}$, and the radio halo turns on shortly after. We
choose this simulation snapshot to analyse. \C{Note that in recent simulations by \citet{miniati15}, the turbulent reacceleration is strongest around core passage. However, we are not very sensitive to the adopted decay time, since the thermal and CR quantities are very similar a few 100 Myrs before and after $z = 0$, where we have chosen to evaluate the simulations.}
In all our calculations we assume that turbulent reacceleration efficiently
accelerates particles for $\tau_{\rm cl} \sim 650$ Myrs (which is roughly the
cascade time on which turbulence is damped) and that during this turbulent phase
CR streaming and spatial diffusion can be neglected. In
Section~\ref{sect:self-limiting}, we explore sensitivity to the last assumption.

@@

The time evolution of the spectral energy distribution of CRps,
$f_{\rmn{p}}$, is similarly given by:
\begin{eqnarray}
\lefteqn{
  {{d f_{\rmn{p}}(p,t)}\over{d t}} =
  {{\partial }\over{\partial p}}
  \left\{
  f_{\rmn{p}}(p,t)\left[ \left|{{d p}\over{d t}}\right|_{\rm Coul}
    + \frac{p}{3}\bnabla\bcdot \left(\bvel+\bvel_{\rmn{st}}\right)\right.\right.}
\nonumber\\
&&-\left.\left. {1\over{p^2}}{{\partial }\over{\partial p}}\left(p^2 D_{pp}\right)
\right]\right\} - f_{\rmn{p}}(p,t) \bnabla\bcdot \left(\bvel+\bvel_{\rmn{st}}\right) 
- \bvel_{\rmn{st}}\bcdot\bnabla f_{\rmn{p}}(p,t)
\nonumber\\
&&+\,\,\, {{\partial^2 }\over{\partial p^2}}
\left[ D_{pp} f_{\rmn{p}}(p,t) \right] - {{f_{\rmn{p}}(p,t)}\over{\tau_{\rm had}(p)}}
+ Q_{\rmn{p}}(p,t)\, ,
\label{eq:FP_p}
\end{eqnarray}
where $\bvel_{\rmn{st}}=-\varv_{\rmn{A}}\bnabla f_{\rmn{p}}/|\bnabla
f_{\rmn{p}}|$ is the streaming velocity in the isotropic transport
approximation, $\varv_{\rmn{A}}$ is the Alfv\'en speed,
and $Q_{\rmn{p}}(p,t)$ denotes the injection rate of shock accelerated
CRps as a function of momentum $p=P/(m_{\rmn{p}} c)$ and time $t$ (see
section~\ref{sec:cosmo_sim}). The timescale of
hadronic losses that produce pions via CRp collisions with thermal
protons of the ICM is given by 
\begin{eqnarray}
  \tau_{\rm had} = \left[c\,n_\rmn{th}\,\sigma^{+/-,0}(p)\right]^{-1}\,,
\end{eqnarray}
where we use the cross-section, $\sigma^{+/-,0}(p)$, given by the
fitting formula in \citet{1986ApJ...307...47D}.

@@

We adopt a Kraichnan spectrum ($s=3/2$) for the fast modes, as seen in simulations \citep{cho03}. This can also be written as: 
\begin{equation}
  \label{eq:Wk}
  \mathcal{W}(k) =
\sqrt{2/7\,I_L\,\rho\,\langle \Vph \rangle}\,k^{-3/2},
\end{equation}
where $I_L$ is the volumetric injection rate of turbulence at scale $L$ (which
is assumed to be constant), $\varv_{\rm ph}\approx c_{\rm s}$ is the phase speed
of waves. \C{Since the Kraichnan spectrum is critical in what follows, it is
  worthwhile taking a moment to recount the origin of the spectral slope $s$ and
  the cascade rate. In the magnetically dominated regime, two counterpropagating
  wave packets of scale $l$ interact on a wave crossing time $\tau_{\rm
    ph}=l/\varv_{\rm ph}$, rather than the eddy turnover time $\tau_{\rm edd} =
  l/\varv_{\rm l}$.  Since $\tau_{\rm ph} \ll \tau_{\rm edd}$, each interaction
  results in a small velocity change $\delta \varv_{\rm l} \sim \varv_{\rm l}
  (\tau_{\rm ph}/\tau_{\rm edd})$. If these changes behave like a random walk,
  the cascade time $\tau_{\rm l}$ it takes for an eddy to become non-linear
  ($\Delta \varv \sim \varv_{\rm l})$ and cascades to smaller scales occurs with
  a characteristic velocity $\varv_{\rm l} \sim \delta \varv_{\rm l} (\tau_{\rm
    l}/\tau_{\rm ph})^{1/2} \sim \varv_{\rm l} (\tau_{\rm ph}/\tau_{\rm edd})\,
  (\tau_{\rm l}/\tau_{\rm ph})^{1/2}$, which we can solve to obtain the cascade
  time:
\begin{equation}
  \label{eqn:tau_l}
\tau_{\rm l} \sim \left(\frac{\tau_{\rm edd}}{\tau_{\rm ph}}\right)^{2} \tau_{\rm ph} \sim \frac{l \varv_{\rm ph}}{\varv_{\rm l}^{2}} 
\end{equation}
where we have implicitly assumed isotropy. This implies a dissipation rate $\epsilon \sim \varv_{\rm l}^{2}/\tau_{\rm l} \sim \varv_{\rm l}^{4}/(l \varv_{\rm ph})$, or $\varv_{\rm l} \sim (\epsilon l \varv_{\rm ph})^{1/4}$. Thus, since $k E(k)\sim \varv_{\rm l}^{2}$, this gives a kinetic energy spectrum: 
\begin{equation}
E(k) \sim (\epsilon \varv_{\rm ph})^{1/2} k^{-3/2}.
\end{equation}}

It is important to realise that this Kraichnan scaling only applies at small scales, where turbulence is magneto-hydrodynamical. At large scales, turbulent motions are subsonic (${\mathcal M}_{\rm s} \sim 0.2-0.5$) but super-Alfv{\'e}nic (${\mathcal M}_{\rm A} \sim 5$). Thus, except in the case where motions are transsonic and weak shocks become important, motions are fundamentally hydrodynamic and turbulence follows a Kolmogorov ($s=5/3$) spectrum. However, while the turbulent energy density $\eps_{\rm turb} \propto l^{2/3}$ decreases at small scales, the magnetic energy density \C{(which is dominated by the large scale mean field)} is scale-independent. Thus, at some scale $l_{\rm A} \sim l_L {\mathcal M}_{\rm A}^{-3}$, where $\eps_{\rm turb} \sim \eps_{B}$, the magnetic field becomes dynamically important, and turbulence transitions to the MHD regime. The cutoff scale in equation (\ref{eqn:Wk}) can be computed by setting the turbulent cascade rate for fast modes to the transit time damping rate on thermal electrons. This yields \citep{brunetti07,miniati15}:
\begin{equation}
k_{\rm cut} \approx A k_{\rmn{outer}} \frac{\langle (\delta \varv_{\rmn{outer}})^{2}\rangle^{2}}{c_{\rm s}^{4}}
\label{eqn:k_c1} 
\end{equation}
where $A \approx 11 000$. \C{If there is an unimpeded cascade from the injection to the cutoff scale, $k_{\rmn{outer}}=k_L$ and $\varv_{\rmn{outer}}=\varv_{\rmn{c}}$. However, in our case the hydrodynamic (Kolmogorov) cascade transitions to the MHD cascade at the Alfv{\'e}n scale so that $k_{\rmn{outer}}=k_{\rmn{A}}$ and $\varv_{\rmn{outer}}=\varv_{\rmn{A}}$ and hence, we get\footnote{\C{Note that this differs from previous work, which assumes Kraichnan turbulence from the injection scale onward and effectively adopts wave number and compressible velocity at the outer scale $L$ for this estimate of $k_{\rm cut}$.}}
\begin{equation}
k_{\rm cut} \approx A k_{\rmn{A}} \beta^{-2}\,.
\label{eqn:k_c} 
\end{equation}
}
This gives $2\upi/k_\rmn{cut}\sim 0.1-1$~kpc in the ICM. This constitutes an
effective mean free path for CRs, unless plasma instabilities can
mediate interactions between turbulence and particles on smaller
scales \citep{brunetti11}, a possibility we discuss in Section~\ref{sect:self-limiting}. Another possibility is that compressible modes dissipate in weak shocks, resulting in Burgers' turbulence $s=2$ \citep{kowal10,porter15,miniati15}. If Burgers turbulence dominates, then particle acceleration rates are too slow in the face of cooling processes to explain radio haloes \citep{miniati15}, and an alternative model for radio haloes is required.

@@

What is a typical energy density in compressive fast modes? The total turbulent
energy is typically $\sim 15-70$ per cent of the thermal energy in a cluster
\citep{vazza11}; it rises rapidly during a merger. The compressible component is
$\sim 20-40$ per cent of the total turbulent energy, and shows more rapid temporal
variations compared to the \C{incompressible} component
\citep{2013ApJ...771..131B,miniati15}. Note the compressible component in
cluster simulations is much larger than in stirring box simulations
with similar Mach numbers \citep{kowal10,lynn14}. This is likely due to the
compressive nature of turbulent driving (transsonic infall and merger), whereas
idealised simulations tend to use incompressible solenoidal driving and allow
compressive fluctuations to develop on their own. Overall, we adopt a
compressive energy density which is $X_{\rm tu} \sim 0.2$ of the thermal energy
as a baseline estimate.
